---
title: "Latent Semantic Psychometrics: Zero-Shot Item Classification"
author: "Nathan Clarke"
date: "2025-02-28"
output: 
  html_document:
    toc: true
    number_sections: true
---

# Background

Zero-shot item classficiation come be done with domain labels (e.g., ...) or with a description of the domain.

A well-designed PRO should have some conceptual domain definition, so if a description is not available, this should already serve as a flag on the semantic alignment and design of the scale (it's challening to produce an item that represents a domain that is not well-defined...)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 20, fig.height = 15, dpi = 150, eval = TRUE)


library(text)       # Primary package for advanced text-based NLP.
library(dplyr)      # Data manipulation
library(ggplot2)    # Visualization
library(tidyr)      # Data restructuring
library(stringr)    # String operations
library(tidytext)   # Text processing
library(viridis)    # Color palettes
library(EGAnet)  # EGA package for network analysis

# 1. CREATE THE COMPLETE TFI DATASET (ALL 25 ITEMS) ----

tfi_data <- data.frame(
  QuestionID = 1:25,
  QuestionText = c(
    # Intrusive (I)
    "What percentage of your time awake were you consciously aware of your tinnitus?",
    "How strong or loud was your tinnitus?",
    "What percentage of your time awake were you annoyed by your tinnitus?",

    # Sense of Control (SC)
    "Did you feel in control in regard to your tinnitus?",
    "How easy was it for you to cope with your tinnitus?",
    "How easy was it for you to ignore your tinnitus?",

    # Cognitive (C)
    "How much did your tinnitus interfere with your ability to concentrate?",
    "How much did your tinnitus interfere with your ability to think clearly?",
    "How much did your tinnitus interfere with your ability to focus attention on other things besides your tinnitus?",

    # Sleep (SL)
    "How often did your tinnitus make it difficult to fall asleep or stay asleep?",
    "How often did your tinnitus cause you difficulty in getting as much sleep as you needed?",
    "How much of the time did your tinnitus keep you from sleeping as deeply or as peacefully as you would have liked?",

    # Auditory (A)
    "How much has your tinnitus interfered with your ability to hear clearly?",
    "How much has your tinnitus interfered with your ability to understand people who are talking?",
    "How much has your tinnitus interfered with your ability to follow conversations in a group or at meetings?",

    # Relaxation (R)
    "How much has your tinnitus interfered with your quiet resting activities?",
    "How much has your tinnitus interfered with your ability to relax?",
    "How much has your tinnitus interfered with your ability to enjoy peace and quiet?",

    # Quality of Life (Q)
    "How much has your tinnitus interfered with your enjoyment of social activities?",
    "How much has your tinnitus interfered with your enjoyment of life?",
    "How much has your tinnitus interfered with your relationships with family, friends, and other people?",
    "How often did your tinnitus cause you difficulty performing your work or other tasks, such as home maintenance, schoolwork, or caring for children or others?",

    # Emotional (E)
    "How anxious or worried has your tinnitus made you feel?",
    "How bothered or upset have you been because of your tinnitus?",
    "How depressed were you because of your tinnitus?"
  ),
  IntendedDomain = c(
    rep("Intrusive", 3),
    rep("Sense of Control", 3),
    rep("Cognitive", 3),
    rep("Sleep", 3),
    rep("Auditory", 3),
    rep("Relaxation", 3),
    rep("Quality of Life", 4),
    rep("Emotional", 3)
  ),
  stringsAsFactors = FALSE
)

candidate_domains <- c(
  "Intrusive", "Sense of Control", "Cognitive", "Sleep",
  "Auditory", "Relaxation", "Quality of Life", "Emotional"
)

tfi_domains <- tibble(
  Domain = c(
    "Auditory", "Cognitive", "Emotional", "Intrusive",
    "Quality of Life", "Relaxation", "Sense of Control", "Sleep"
  ),
  Definition = c(
    "Evaluates problems hearing clearly due to tinnitus.",
    "Captures difficulties in concentration and cognitive tasks due to tinnitus.",
    "Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression).",
    "Measures how much the tinnitus intrudes on consciousness and daily life.",
    "Measures the effect of tinnitus on overall enjoyment and engagement in life activities.",
    "Examines the impact of tinnitus on relaxation and quiet activities.",
    "Determines how much control the patient feels they have over their tinnitus.",
    "Assesses the extent to which tinnitus interferes with sleep."
  )
)
```

# Zero-shot classification OF TFI ITEMS ACCORDING TO DOMAIN LABELS (e.g. "Intrusive")
## Single-label classification

```{r}

# 2a. ZERO-SHOT CLASSIFICATION OF TFI ITEMS ACCORDING TO DOMAIN LABELS (e.g. "Intrusive") (multi_label = F) ----

# classification_results <- data.frame(
#   Item = character(),
#   Predicted_Domain = character(),
#   Confidence = numeric(),
#   stringsAsFactors = FALSE
# )
#
# for (i in 1:length(tfi_data$QuestionText)) {
#   test_result <- textZeroShot(tfi_data$QuestionText[i], candidate_domains)
#   test_result$Item <- tfi_data$QuestionText[i]
#   classification_results <- rbind(classification_results, test_result)
# }


# 2b. DATA TRANSFORMATION: RESTRUCTURING OUTPUT ----

# # Pivot labels into long format, keeping Item for context
# classification_results_long <- classification_results %>%
#   pivot_longer(
#     cols = starts_with("labels_x"),
#     names_to = "Label_Index",
#     values_to = "Predicted_Domain"
#   ) %>%
#   mutate(Score_Index = gsub("labels", "scores", Label_Index)) %>%
#   select(sequence, Item, Label_Index, Predicted_Domain, Score_Index) # Retain Score_Index for merging
#
# # Pivot scores separately
# classification_scores_long <- classification_results %>%
#   pivot_longer(
#     cols = starts_with("scores_x_"),
#     names_to = "Score_Index",
#     values_to = "Confidence"
#   ) %>%
#   select(sequence, Score_Index, Confidence) # Keep only necessary columns
#
#
# # Perform the join using both sequence and Score_Index
# classification_results_long <- classification_results_long %>%
#   left_join(classification_scores_long, by = c("sequence", "Score_Index")) %>%
#   arrange(sequence, -Confidence) %>%
#   select(Item, Predicted_Domain, Confidence)
#
# # Using the tfi_data for reference, add question ID to classification_results_long and arrange by QuestionID
# classification_results_long <- classification_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText")) %>%
#   select(QuestionID, Item, Predicted_Domain, Confidence) %>%
#   arrange(QuestionID)
#
# # Using the tfi_data for reference, add predicted domain to classification_results_long
# classification_results_long <- classification_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText"))
#
# #drop rename to QuestionID.x to QuestionID, drop QuestionID.y, and select item, questionID, item, intended domain, predicted domain, and confidence
# classification_results_long <- classification_results_long %>%
#   rename(QuestionID = QuestionID.x) %>%
#   select(QuestionID, Item, IntendedDomain, Predicted_Domain, Confidence)

# write.csv(classification_results_long, "TFI_ZeroShot_Classifications_Long.csv", row.names = FALSE)

```

```{r}
# read saved results
classification_results_long <- read.csv("TFI_ZeroShot_Classifications_Long.csv")

# 2c. VISUALIZATION: INDIVIDUAL ITEM PLOTS ----

for (item in unique(classification_results_long$Item)) {
  item_data <- classification_results_long %>% filter(Item == item)
  item_number <- unique(item_data$QuestionID)
  intended_domain <- unique(item_data$IntendedDomain)

  p <- ggplot(item_data, aes(
    x = reorder(Predicted_Domain, -Confidence),
    y = Confidence,
    fill = Predicted_Domain
  )) +
    geom_bar(stat = "identity", width = 0.6) +
    theme_minimal() +
    labs(
      title = paste("Item", item_number, ":", item),
      subtitle = paste("Intended Domain:", intended_domain),
      x = "Predicted Domain",
      y = "Confidence Score",
      fill = "Predicted Domain"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      legend.position = "bottom"
    )

  print(p)
}

# Using classification_results_long, summarise how many items were correctly classified (i.e., the largest confidence score was also it's predicted domain)
correctly_classified <- classification_results_long %>%
  group_by(QuestionID) %>%
  filter(Confidence == max(Confidence)) %>%
  filter(Predicted_Domain == IntendedDomain) %>%
  summarise(Correctly_Classified = n())

```

## Multi-label classification

```{r}
# 3a. ZERO-SHOT CLASSIFICATION OF TFI ITEMS ACCORDING TO DOMAIN LABELS (e.g. "Intrusive") (multi_label = T) ----

# classification_results_multi <- data.frame(
#   Item = character(),
#   Predicted_Domain = character(),
#   Confidence = numeric(),
#   stringsAsFactors = FALSE
# )
#
# for (i in 1:length(tfi_data$QuestionText)) {
#   test_result <- textZeroShot(tfi_data$QuestionText[i], candidate_domains, multi_label = TRUE)
#   test_result$Item <- tfi_data$QuestionText[i]
#   classification_results_multi <- rbind(classification_results_multi, test_result)
# }

# 3b. DATA TRANSFORMATION: RESTRUCTURING OUTPUT ----

# # Pivot labels into long format, keeping Item for context
# classification_results_multi_long <- classification_results_multi %>%
#   pivot_longer(
#     cols = starts_with("labels_x"),
#     names_to = "Label_Index",
#     values_to = "Predicted_Domain"
#   ) %>%
#   mutate(Score_Index = gsub("labels", "scores", Label_Index)) %>%
#   select(sequence, Item, Label_Index, Predicted_Domain, Score_Index) # Retain Score_Index for merging
#
# # Pivot scores separately
# classification_scores_multi_long <- classification_results_multi %>%
#   pivot_longer(
#     cols = starts_with("scores_x_"),
#     names_to = "Score_Index",
#     values_to = "Confidence"
#   ) %>%
#   select(sequence, Score_Index, Confidence) # Keep only necessary columns
#
#
# # Perform the join using both sequence and Score_Index
# classification_results_multi_long <- classification_results_multi_long %>%
#   left_join(classification_scores_multi_long, by = c("sequence", "Score_Index")) %>%
#   arrange(sequence, -Confidence) %>%
#   select(Item, Predicted_Domain, Confidence)
#
# # Using the tfi_data for reference, add question ID to classification_results_long and arrange by QuestionID
# classification_results_multi_long <- classification_results_multi_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText")) %>%
#   select(QuestionID, Item, Predicted_Domain, Confidence) %>%
#   arrange(QuestionID)
#
# # Using the tfi_data for reference, add predicted domain to classification_results_long
# classification_results_multi_long <- classification_results_multi_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText"))
#
# #drop rename to QuestionID.x to QuestionID, drop QuestionID.y, and select item, questionID, item, intended domain, predicted domain, and confidence
# classification_results_multi_long <- classification_results_multi_long %>%
#   rename(QuestionID = QuestionID.x) %>%
#   select(QuestionID, Item, IntendedDomain, Predicted_Domain, Confidence)
#
# #write.csv(classification_results_multi_long, "TFI_ZeroShot_Classifications_Multi_Long.csv", row.names = FALSE)
```

```{r}
# read saved results
classification_results_multi_long <- read.csv("TFI_ZeroShot_Classifications_Multi_Long.csv")

# 3c. VISUALIZATION: INDIVIDUAL ITEM PLOTS ----

for (item in unique(classification_results_multi_long$Item)) {
  item_data <- classification_results_multi_long %>% filter(Item == item)
  item_number <- unique(item_data$QuestionID)
  intended_domain <- unique(item_data$IntendedDomain)

  p <- ggplot(item_data, aes(
    x = reorder(Predicted_Domain, -Confidence),
    y = Confidence,
    fill = Predicted_Domain
  )) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_hline(yintercept = 0.7, linetype = "dashed", color = "darkgrey", linewidth = 1) + # Dark grey threshold line
    annotate("text",
      x = length(unique(item_data$Predicted_Domain)) + 0.2, # Further right
      y = 0.64, # Lowered text position
      label = "High Confidence\nThreshold (0.7)", # Line break added
      color = "darkgrey", hjust = 1, size = 3.2
    ) + # Reduced text size
    theme_minimal() +
    labs(
      title = paste("Item", item_number, ":", item),
      subtitle = paste("Intended Domain:", intended_domain),
      x = "Predicted Domain",
      y = "Confidence Score",
      fill = "Predicted Domain"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      legend.position = "bottom"
    )

  print(p)
}
```

```{r}
# Tabular summaries
# Using classification_results_multi_long, summarise how many items were correctly classified (i.e., the largest confidence score was also it's predicted domain)
correctly_classified_multi <- classification_results_multi_long %>%
  group_by(QuestionID) %>%
  filter(Confidence == max(Confidence)) %>%
  filter(Predicted_Domain == IntendedDomain) %>%
  summarise(Correctly_Classified = n())

# Using classification_results_multi_long, summarise how many items had confidence scores >0.7
high_confidence_multi <- classification_results_multi_long %>%
  group_by(QuestionID) %>%
  filter(Confidence > 0.7) %>%
  summarise(High_Confidence = n())

# Using classification_results_multi_long, filter to all rows where confidence scores >0.7 and where the predicted domain does not match the intended domain
high_confidence_incorrect_multi <- classification_results_multi_long %>%
  filter(Confidence > 0.7) %>%
  filter(Predicted_Domain != IntendedDomain)
```

# ZERO-SHOT CLASSIFICATION OF TFI ITEMS ACCORDING TO DOMAIN DESCRIPTION in MEIKLE et al. 2012
## Multi-label

```{r}
# # 
# # 6a. ZERO-SHOT CLASSIFICATION OF TFI ITEMS ACCORDING TO DOMAIN DESCRIPTION in MEIKLE et al. 2012) (multi_label = T) ----
# # 
#
# # zeroshot_results <- data.frame(
# #   Item = character(),
# #   Predicted_Domain = character(),
# #   Confidence = numeric(),
# #   stringsAsFactors = FALSE
# # )
# #
# # for (i in 1:length(tfi_data$QuestionText)) {
# #   test_result <- textZeroShot(tfi_data$QuestionText[i], tfi_domains$Definition, multi_label = TRUE)
# #   test_result$Item <- tfi_data$QuestionText[i]
# #   zeroshot_results <- rbind(zeroshot_results, test_result)
# # }
```

```{r}
 # 
# # 6b. DATA TRANSFORMATION: RESTRUCTURING OUTPUT ----
# 
#
#
# # Pivot labels into long format, keeping Item for context
# zeroshot_results_long <- zeroshot_results %>%
#   pivot_longer(
#     cols = starts_with("labels_x"),
#     names_to = "Label_Index",
#     values_to = "Predicted_Domain"
#   ) %>%
#   mutate(Score_Index = gsub("labels", "scores", Label_Index)) %>%
#   select(sequence, Item, Label_Index, Predicted_Domain, Score_Index) # Retain Score_Index for merging
#
# # Pivot scores separately
# zeroshot_scores_multi_long <- zeroshot_results %>%
#   pivot_longer(
#     cols = starts_with("scores_x_"),
#     names_to = "Score_Index",
#     values_to = "Confidence"
#   ) %>%
#   select(sequence, Score_Index, Confidence) # Keep only necessary columns
#
#
# # Perform the join using both sequence and Score_Index
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(zeroshot_scores_multi_long, by = c("sequence", "Score_Index")) %>%
#   arrange(sequence, -Confidence) %>%
#   select(Item, Predicted_Domain, Confidence)
#
# # Using the tfi_data for reference, add question ID to classification_results_long and arrange by QuestionID
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText")) %>%
#   select(QuestionID, Item, Predicted_Domain, Confidence) %>%
#   arrange(QuestionID)
#
# # Using the tfi_data for reference, add predicted domain to classification_results_long
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText"))
#
# #drop rename to QuestionID.x to QuestionID, drop QuestionID.y, and select item, questionID, item, intended domain, predicted domain, and confidence
# zeroshot_results_long <- zeroshot_results_long %>%
#   rename(QuestionID = QuestionID.x) %>%
#   select(QuestionID, Item, IntendedDomain, Predicted_Domain, Confidence)
#
# # ✅ Add Intended and Predicted Domain Definitions (Label + Definition)
# zeroshot_results_long <- zeroshot_results_long %>%
#   mutate(
#     # Intended Domain Definitions
#     Intended_Domain_Definition = case_when(
#       IntendedDomain == "Auditory" ~ "Auditory: Evaluates problems hearing clearly due to tinnitus.",
#       IntendedDomain == "Cognitive" ~ "Cognitive: Captures difficulties in concentration and cognitive tasks due to tinnitus.",
#       IntendedDomain == "Emotional" ~ "Emotional: Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression).",
#       IntendedDomain == "Intrusive" ~ "Intrusive: Measures how much the tinnitus intrudes on consciousness and daily life.",
#       IntendedDomain == "Quality of Life" ~ "Quality of Life: Measures the effect of tinnitus on overall enjoyment and engagement in life activities.",
#       IntendedDomain == "Relaxation" ~ "Relaxation: Examines the impact of tinnitus on relaxation and quiet activities.",
#       IntendedDomain == "Sense of Control" ~ "Sense of Control: Determines how much control the patient feels they have over their tinnitus.",
#       IntendedDomain == "Sleep" ~ "Sleep: Assesses the extent to which tinnitus interferes with sleep.",
#       TRUE ~ NA_character_
#     ),
#
#     # Predicted Domain Definitions
#     Predicted_Domain_Definition = case_when(
#       Predicted_Domain == "Evaluates problems hearing clearly due to tinnitus." ~ "Auditory: Evaluates problems hearing clearly due to tinnitus.",
#       Predicted_Domain == "Captures difficulties in concentration and cognitive tasks due to tinnitus." ~ "Cognitive: Captures difficulties in concentration and cognitive tasks due to tinnitus.",
#       Predicted_Domain == "Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression)." ~ "Emotional: Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression).",
#       Predicted_Domain == "Measures how much the tinnitus intrudes on consciousness and daily life." ~ "Intrusive: Measures how much the tinnitus intrudes on consciousness and daily life.",
#       Predicted_Domain == "Measures the effect of tinnitus on overall enjoyment and engagement in life activities." ~ "Quality of Life: Measures the effect of tinnitus on overall enjoyment and engagement in life activities.",
#       Predicted_Domain == "Examines the impact of tinnitus on relaxation and quiet activities." ~ "Relaxation: Examines the impact of tinnitus on relaxation and quiet activities.",
#       Predicted_Domain == "Determines how much control the patient feels they have over their tinnitus." ~ "Sense of Control: Determines how much control the patient feels they have over their tinnitus.",
#       Predicted_Domain == "Assesses the extent to which tinnitus interferes with sleep." ~ "Sleep: Assesses the extent to which tinnitus interferes with sleep.",
#       TRUE ~ NA_character_
#     )
#   )


# write.csv(zeroshot_results_long, "TFI_zeroshot_domains_results_multilabel.csv", row.names = FALSE)
```

```{r}
# ==================================================
# 6c. VISUALIZATION: INDIVIDUAL ITEM PLOTS (FLIPPED & REFINED) ----
# ==================================================

zeroshot_results_long <- read.csv("TFI_zeroshot_domains_results_multilabel.csv")

for (item in unique(zeroshot_results_long$Item)) {
  # ✅ Filter item data
  item_data <- zeroshot_results_long %>% filter(Item == item)
  item_number <- unique(item_data$QuestionID) # ✅ Corrected
  intended_domain <- unique(item_data$IntendedDomain) # ✅ Corrected

  # ✅ Wrap the title for better readability
  wrapped_title <- str_wrap(paste("Zero-Shot Classification (multi-label) for Item", item_number, ":", item), width = 80)

  # ✅ Generate flipped plot
  p <- ggplot(item_data, aes(
    y = reorder(Predicted_Domain_Definition, Confidence), # ✅ Flip to y-axis
    x = Confidence
  )) +
    geom_bar(stat = "identity", width = 0.6) +

    # ✅ Convert high-confidence threshold into a vertical reference line
    geom_vline(xintercept = 0.7, linetype = "dashed", color = "darkgrey", linewidth = 1) +

    # ✅ Position threshold label in the bottom right of each plot
    annotate("text",
      y = 0.8, # ✅ Near the bottom of the y-axis
      x = 0.72, # ✅ Slightly past the reference line
      label = "High Confidence\nThreshold (0.7)",
      color = "darkgrey", hjust = 0, size = 3.2
    ) +
    theme_minimal() +

    # ✅ Adjust confidence score scale with 0.05 increments
    scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.05)) +
    labs(
      title = wrapped_title, # ✅ Wrapped title for better readability
      subtitle = paste("Intended Domain:", intended_domain),
      x = "Confidence Score",
      y = "Predicted Domain Definition" # ✅ Adjusted for flipped axis
    ) +
    theme(
      axis.text.y = element_text(size = 12), # ✅ Increase readability for long domain text
      axis.text.x = element_text(size = 12, angle = 45, hjust = 1), # ✅ Rotate x-axis labels
      legend.position = "none" # ✅ Removed unnecessary legend
    )

  print(p)
}
```

## Single-label

```{r}
# 7a. ZERO-SHOT CLASSIFICATION OF TFI ITEMS ACCORDING TO DOMAIN DESCRIPTION in MEIKLE et al. 2012) (multi_label = F) ----


# zeroshot_results <- data.frame(
#   Item = character(),
#   Predicted_Domain = character(),
#   Confidence = numeric(),
#   stringsAsFactors = FALSE
# )
#
# for (i in 1:length(tfi_data$QuestionText)) {
#   test_result <- textZeroShot(tfi_data$QuestionText[i], tfi_domains$Definition, multi_label = FALSE)
#   test_result$Item <- tfi_data$QuestionText[i]
#   zeroshot_results <- rbind(zeroshot_results, test_result)
# }
```

```{r}
# ==================================================
# 7b. DATA TRANSFORMATION: RESTRUCTURING OUTPUT ----
# ==================================================


# # Pivot labels into long format, keeping Item for context
# zeroshot_results_long <- zeroshot_results %>%
#   pivot_longer(
#     cols = starts_with("labels_x"),
#     names_to = "Label_Index",
#     values_to = "Predicted_Domain"
#   ) %>%
#   mutate(Score_Index = gsub("labels", "scores", Label_Index)) %>%
#   select(sequence, Item, Label_Index, Predicted_Domain, Score_Index) # Retain Score_Index for merging
#
# # Pivot scores separately
# zeroshot_scores_multi_long <- zeroshot_results %>%
#   pivot_longer(
#     cols = starts_with("scores_x_"),
#     names_to = "Score_Index",
#     values_to = "Confidence"
#   ) %>%
#   select(sequence, Score_Index, Confidence) # Keep only necessary columns
#
#
# # Perform the join using both sequence and Score_Index
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(zeroshot_scores_multi_long, by = c("sequence", "Score_Index")) %>%
#   arrange(sequence, -Confidence) %>%
#   select(Item, Predicted_Domain, Confidence)
#
# # Using the tfi_data for reference, add question ID to classification_results_long and arrange by QuestionID
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText")) %>%
#   select(QuestionID, Item, Predicted_Domain, Confidence) %>%
#   arrange(QuestionID)
#
# # Using the tfi_data for reference, add predicted domain to classification_results_long
# zeroshot_results_long <- zeroshot_results_long %>%
#   left_join(tfi_data, by = c("Item" = "QuestionText"))
#
# #drop rename to QuestionID.x to QuestionID, drop QuestionID.y, and select item, questionID, item, intended domain, predicted domain, and confidence
# zeroshot_results_long <- zeroshot_results_long %>%
#   rename(QuestionID = QuestionID.x) %>%
#   select(QuestionID, Item, IntendedDomain, Predicted_Domain, Confidence)
#
# # ✅ Add Intended and Predicted Domain Definitions (Label + Definition)
# zeroshot_results_long <- zeroshot_results_long %>%
#   mutate(
#     # Intended Domain Definitions
#     Intended_Domain_Definition = case_when(
#       IntendedDomain == "Auditory" ~ "Auditory: Evaluates problems hearing clearly due to tinnitus.",
#       IntendedDomain == "Cognitive" ~ "Cognitive: Captures difficulties in concentration and cognitive tasks due to tinnitus.",
#       IntendedDomain == "Emotional" ~ "Emotional: Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression).",
#       IntendedDomain == "Intrusive" ~ "Intrusive: Measures how much the tinnitus intrudes on consciousness and daily life.",
#       IntendedDomain == "Quality of Life" ~ "Quality of Life: Measures the effect of tinnitus on overall enjoyment and engagement in life activities.",
#       IntendedDomain == "Relaxation" ~ "Relaxation: Examines the impact of tinnitus on relaxation and quiet activities.",
#       IntendedDomain == "Sense of Control" ~ "Sense of Control: Determines how much control the patient feels they have over their tinnitus.",
#       IntendedDomain == "Sleep" ~ "Sleep: Assesses the extent to which tinnitus interferes with sleep.",
#       TRUE ~ NA_character_
#     ),
#
#     # Predicted Domain Definitions
#     Predicted_Domain_Definition = case_when(
#       Predicted_Domain == "Evaluates problems hearing clearly due to tinnitus." ~ "Auditory: Evaluates problems hearing clearly due to tinnitus.",
#       Predicted_Domain == "Captures difficulties in concentration and cognitive tasks due to tinnitus." ~ "Cognitive: Captures difficulties in concentration and cognitive tasks due to tinnitus.",
#       Predicted_Domain == "Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression)." ~ "Emotional: Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression).",
#       Predicted_Domain == "Measures how much the tinnitus intrudes on consciousness and daily life." ~ "Intrusive: Measures how much the tinnitus intrudes on consciousness and daily life.",
#       Predicted_Domain == "Measures the effect of tinnitus on overall enjoyment and engagement in life activities." ~ "Quality of Life: Measures the effect of tinnitus on overall enjoyment and engagement in life activities.",
#       Predicted_Domain == "Examines the impact of tinnitus on relaxation and quiet activities." ~ "Relaxation: Examines the impact of tinnitus on relaxation and quiet activities.",
#       Predicted_Domain == "Determines how much control the patient feels they have over their tinnitus." ~ "Sense of Control: Determines how much control the patient feels they have over their tinnitus.",
#       Predicted_Domain == "Assesses the extent to which tinnitus interferes with sleep." ~ "Sleep: Assesses the extent to which tinnitus interferes with sleep.",
#       TRUE ~ NA_character_
#     )
#   )


# write.csv(zeroshot_results_long, "TFI_zeroshot_domains_results_singlelabel.csv", row.names = FALSE)
```

```{r}
# ==================================================
# 7c. VISUALIZATION: INDIVIDUAL ITEM PLOTS (FLIPPED & REFINED) ----
# ==================================================

zeroshot_results_long <- read.csv("TFI_zeroshot_domains_results_singlelabel.csv")

for (item in unique(zeroshot_results_long$Item)) {
  # ✅ Filter item data
  item_data <- zeroshot_results_long %>% filter(Item == item)
  item_number <- unique(item_data$QuestionID) # ✅ Corrected
  intended_domain <- unique(item_data$IntendedDomain) # ✅ Corrected

  # ✅ Wrap the title for better readability
  wrapped_title <- str_wrap(paste("Zero-Shot Classification (single-label) for Item", item_number, ":", item), width = 80)

  # ✅ Generate flipped plot
  p <- ggplot(item_data, aes(
    y = reorder(Predicted_Domain_Definition, Confidence), # ✅ Flip to y-axis
    x = Confidence
  )) +
    geom_bar(stat = "identity", width = 0.6) +

    # ✅ Convert high-confidence threshold into a vertical reference line
    geom_vline(xintercept = 0.7, linetype = "dashed", color = "darkgrey", linewidth = 1) +

    # ✅ Position threshold label in the bottom right of each plot
    annotate("text",
      y = 0.8, # ✅ Near the bottom of the y-axis
      x = 0.72, # ✅ Slightly past the reference line
      label = "High Confidence\nThreshold (0.7)",
      color = "darkgrey", hjust = 0, size = 3.2
    ) +
    theme_minimal() +

    # ✅ Adjust confidence score scale with 0.05 increments
    scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.05)) +
    labs(
      title = wrapped_title, # ✅ Wrapped title for better readability
      subtitle = paste("Intended Domain:", intended_domain),
      x = "Confidence Score",
      y = "Predicted Domain Definition" # ✅ Adjusted for flipped axis
    ) +
    theme(
      axis.text.y = element_text(size = 12), # ✅ Increase readability for long domain text
      axis.text.x = element_text(size = 12, angle = 45, hjust = 1), # ✅ Rotate x-axis labels
      legend.position = "none" # ✅ Removed unnecessary legend
    )

  print(p)
}
```

```{r}
zeroshot_results_long <- zeroshot_results_long %>%
  mutate(
    # Predicted Domain Definitions
    predicted_domain_label = case_when(
      Predicted_Domain == "Evaluates problems hearing clearly due to tinnitus." ~ "Auditory",
      Predicted_Domain == "Captures difficulties in concentration and cognitive tasks due to tinnitus." ~ "Cognitive",
      Predicted_Domain == "Evaluates the emotional distress caused by tinnitus (e.g., anxiety, frustration, depression)." ~ "Emotional",
      Predicted_Domain == "Measures how much the tinnitus intrudes on consciousness and daily life." ~ "Intrusive",
      Predicted_Domain == "Measures the effect of tinnitus on overall enjoyment and engagement in life activities." ~ "Quality of Life",
      Predicted_Domain == "Examines the impact of tinnitus on relaxation and quiet activities." ~ "Relaxation",
      Predicted_Domain == "Determines how much control the patient feels they have over their tinnitus." ~ "Sense of Control",
      Predicted_Domain == "Assesses the extent to which tinnitus interferes with sleep." ~ "Sleep",
      TRUE ~ NA_character_
    )
  )

# select the largest confidence score for each item as a new dataframe
largest_confidence <- zeroshot_results_long %>%
  group_by(QuestionID) %>%
  filter(Confidence == max(Confidence)) %>%
  select(QuestionID, predicted_domain_label, Confidence)

# summarise the each predicted_domain_label frequency and arrange by descending frequency
domain_frequency <- largest_confidence %>%
  group_by(predicted_domain_label) %>%
  summarise(Frequency = n()) %>%
  arrange(desc(Frequency))
```